Ecological management of stochastic systems with long transients

### 2020-01-03 manuscript/manuscript.Rmd
a few captions

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = 1,  
                      dev = "cairo_pdf", fig.width=4.5, 
                      fig.height=3.5, message = FALSE,
                      warning = FALSE)
library(hrbrthemes)
library(Cairo)
library(extrafont)
extrafont::loadfonts()

@@

```{r figure2}
# Fig 2
fig1_sims %>%
  ggplot(aes(time, state, group = interaction(rep, scenario), col= scenario)) + 
  geom_line(alpha=0.3) + scale_color_solarized() + facet_wrap(~scenario)
```

@@

```{r figure4}
# Fig 4
fixed_sims <- read_csv("manuscript_cache/fixed_sims.csv")

@@

Several caveats about this analysis must be noted.  First, I underscore that this is not intended to be a precise model of bark beetle dynamics, but rather a general model  [sensu @Levins1966] to explore the management consequences of confusing a ghost attractor for an alternative stable state.  A precise model would depend on ecological details of forest and beetle dynamics (including spatially explicit dynamics), as well as the details of management interventions and associated costs.  Second, for simplicity, we have assumed that while our manager does not know whether the system has multiple equilibria or a ghost attractor, the manager does know most other details about the system.  Importantly, we have assumed the manager observes the current population size without error before each action.  In reality, any such observation would be accompanied by some uncertainty, making the decision problem into a Partially Observed Markov Decision Process (POMDP), increasing the complexity dramatically [e.g. @pomdp-intro].  I have also assumed the manager has knowledge of the general model structure.  This assumption can be relaxed in much the same way in which introduced model uncertainty, allowing one to consider an arbitrary suite of models (i.e. varying the parameter $a$ over a range of values, as well as to models with completely different structure).  However, the curse of dimensionality imposes obvious limitations on this.  Moreover, the approach assumes that while the ecological dynamics can be transient, the environment in which they occur is stationary (i.e. model parameters are constant.)  It is worth noting that these all of these issues in Markov Decision Processes are active areas of research in the robotics and artificial intelligence community, and recent advances in reinforcement learning could provide a promising way forward for many ecological management problems as well [e.g. @pnas].

>>>>

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = 1,  
                      dev = "cairo_pdf", 
                      fig.width=4.5, 
                      fig.height=3.5, 
                      fig.align = "center",
                      message = FALSE,
                      warning = FALSE)
library(hrbrthemes)
library(Cairo)
library(extrafont)
extrafont::loadfonts()

@@

```{r figure2, fig.cap="50 replicate simulations with identical starting conditions ($X_0 = 21$) of each scenario with parameter values as given in Fig 1. While the ensemble of 50 replicates clearly appears more stable than the ghost attractor, individual replicates may be difficult to distingush."}
# Fig 2
fig1_sims %>%
  ggplot(aes(time, state, group = interaction(rep, scenario), col = scenario)) + 
  geom_line(alpha=0.3) + scale_color_solarized() + facet_wrap(~scenario, ncol = 1)
```

@@

```{r figure4, fig.cap = "Simulations of dynamic management"}
# Fig 4
fixed_sims <- read_csv("manuscript_cache/fixed_sims.csv")

@@

Several caveats about this analysis must be noted.  First, I underscore that this is not intended to be a precise model of bark beetle dynamics, but rather a general model  [sensu @Levins1966] to explore the management consequences of confusing a ghost attractor for an alternative stable state.  A precise model would depend on ecological details of forest and beetle dynamics (including spatially explicit dynamics), as well as the details of management interventions and associated costs.  Second, for simplicity, we have assumed that while our manager does not know whether the system has multiple equilibria or a ghost attractor, the manager does know most other details about the system.  Importantly, we have assumed the manager observes the current population size without error before each action.  In reality, any such observation would be accompanied by some uncertainty, making the decision problem into a Partially Observed Markov Decision Process (POMDP), increasing the complexity dramatically [e.g. @pomdp-intro].  I have also assumed the manager has knowledge of the general model structure.  This assumption can be relaxed in much the same way in which introduced model uncertainty, allowing one to consider an arbitrary suite of models (i.e. varying the parameter $a$ over a range of values, as well as to models with completely different structure).  However, the curse of dimensionality imposes obvious limitations on this.  Moreover, the approach assumes that while the ecological dynamics can be transient, the environment in which they occur is stationary (i.e. model parameters are constant.)  It is worth noting that these all of these issues in Markov Decision Processes are active areas of research in the robotics and artificial intelligence community, and recent advances in reinforcement learning could provide a promising way forward for many ecological management problems as well [e.g. @pnas; or see @deep-rl for an overview].
